# Missing values

---
title: "missing-value"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load library

```{r}
library(tidyverse)
library(extracat)
```

### Load data 

```{r}
index = c('loan_amnt','issue_d','annual_inc','addr_state','grade','purpose','loan_status','home_ownership','emp_length')
df <- read.csv(file="loan.csv", header=TRUE, sep=",")[index]
df[df==""] <- NA
df[df=="n/a"] <- NA
```

### Missing column pattern graph and analysis

```{r}
visna(df, sort = 'b')
```

This is the graph that describes missing column pattern. It shows that we have some missing values in variables 'emp_length' and 'annual_income'. The missing pattern shows that:  
`1. Most of missing values is in the variable 'emp_length'.  
`2. There is no record that have missing values in these two variables at the same time.   
`3. The missing values take only a small (about 5%) part in the total number of records.   
In our data processing and analysis part, we will simply delete those records with missing values.
