---
author:
title: "EDAV Fall 2019 PSet 5, part A"
output: pdf_document

---
```{r, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = TRUE)
```

This assignment is designed to help you get started on the final project. Be sure to review the final project instructions (https://edav.info/project.html), in particular the new section on reproducible workflow (https://edav.info/project.html#reproducible-workflow), which summarizes principles that we've discussed in class.
    
### 1. The Team

[2 points]

a) Who's on the team? (Include names and UNIs)


b) How do you plan to divide up the work? (Grading is on a group basis. The point of asking is to encourage you to think about this.)


### 2. The Questions

[6 points]

List three questions that you hope you will be able to answer from your research.

a)

b)

c)

### 3. Which output format do you plan to use to submit the project? 

[2 points]

(You don't have to use the same format for this assignment -- PSet 5, part A -- and the final project itself.)

Choices are:

pdf_document  

html_document  

bookdown book: https://bookdown.org/yihui/bookdown/

shiny app: https://shiny.rstudio.com/  


(Remember that it's ok to have pieces of the project that don't fit into the chosen output format; in those cases you can provide links to the relevant material.)

### 4. The Data

What is your data source?  What is your method for importing data? Please be specific. Provide relevant information such as any obstacles you're encountering and what you plan to do to overcome them.

[5 points]


### 5. Provide a short summary, including 2-3 graphs, of your initial investigations. 

LendingClub, Corp LC is the first and largest online Peer-to-Peer (“P2P”) platform to facilitate lending and borrowing of unsecured loans ranging from $1,000 to $35,000. Aiming at providing lower cost transaction fees than other financial intermediaries. We will analysis lending club data(from 2007 to 2015) and from 3 different angles (geo, time, stats) to answer three questions mentioned at question 2. The data set is very rich contans 890k obsvervations, including loan information, address information, and collections among others. The following 3 figures are only a peek of the data from location, amount and grade level. 

```{r}
library(ggplot2)
```

```{r}
df <- read.csv(file="loan.csv", header=TRUE, sep=",")
```

```{r}
ggplot(df, aes(x=addr_state, y=loan_amnt, fill=addr_state)) +
  stat_summary(fun.y="sum", geom="bar") + coord_flip()+
  labs(y ="Total Loan Amount",x="State",title="Bar Chart of loan amount for each state")
```


```{r}
ggplot(data=df,aes(loan_amnt, fill=grade))+
  geom_density(alpha=0.25) + 
  facet_grid(grade ~ .)
```
```{r}
ggplot(df, aes(x=grade, y=loan_amnt, fill=grade)) +
  stat_summary(fun.y="sum", geom="bar") +
  labs(y ="Total Loan Amount",title="Total loan amount based on loan grade")
```
